---
title: "DATA 604 HW 2"
author: "Ahsanul Choudhury"
date: "September 10, 2017"
output:
  html_document:
     theme: cerulean
     toc: true
     toc_float:
       collapsed: false
       smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For an $M/M/1$ queue with mean interarrival time 1.25 minutes and mean service time 1 minute, find all the five of $W_q,\: W,\: L_q,\: L,\: and \: \rho$. For each, interpret in words. Be sure to state all of your units (always!), and the relevant time frame of operation.


### 2.6.1


```{r 2.6.1, echo=TRUE}

m_int <- 1.25                 # Mean interval Time
m_serv <- 1                   # Mean Service Time
c <- 1                        # Number of server($M/M/1$ queue)

lambda <- 1 / m_int           # Mean rate of arrival (lambda)
mu <- 1 / m_serv              # Mean service rate (mu)

rho <- lambda / mu            # Utilization of the server

W_q <- rho / (mu * (1 - rho)) # Mean wait in the queue

W <- 1 / (mu * (1 - rho))     # Mean wait in the system

L_q <- rho^2 / (1 - rho)      # Mean number of customers in the queue

L <- rho / (1 - rho)          # Mean number of customers in the system
```


$W_q$ = `r W_q`

$W$ = `r W`

$L_q$ = `r L_q`

$L$ = `r L`

$\rho$ = `r rho`


***


### 2.6.2


```{r 2.6.2, echo=TRUE}
a <- 0.1
b<- 1.9

sd <- sqrt((b - a)^2 / 12)

var <- sd^2   # Variance of service time
m_serv <- (a + b) / 2
mu <- 1 / m_serv

rho <- lambda / mu
W_q <- (lambda * (var + 1 / mu^2)) / (2 * (1 - lambda/mu))
W <- W_q + m_serv
L_q <- lambda * W_q
L <- lambda * W
```

$W_q$ = `r W_q`

$W$ = `r W`

$L_q$ = `r L_q`

$L$ = `r L`

$\rho$ = `r rho`


With uniform distribution the numbers in both system and queue has dropped. Less wait time for service.


***


### 2.6.3


```{r 2.6.3, echo=TRUE}
m <- 1
sd <- sqrt((a^2 + m^2 + b^2 -a*m -a*b - b*m) / 18)

var <- sd^2
m_serv <- (a + m + b) / 3

rho <- lambda / mu
W_q <- (lambda * (var + 1 / mu^2)) / (2 * (1 - lambda/mu))
W <- W_q + m_serv
L_q <- lambda * W_q
L <- lambda * W
```

$W_q$ = `r W_q`

$W$ = `r W`

$L_q$ = `r L_q`

$L$ = `r L`

$\rho$ = `r rho`

Triangular method achieves the lowest numbers in both system and queue. The wait time is the least out of all three.



***


### 2.6.5


```{r 2.6.5, echo=TRUE}
m_int <- 1.25                 
m_serv <- 3                  
c <- 3                        # Number of server(M/M/3 queue)

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho)))  # Probability that there are 0 customers in the system

lambda <- 1 / m_int           
mu <- 1 / m_serv              

rho <- lambda / (c * mu)            

L_q <- (p0 * (lambda / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))
L <- L_q + lambda / mu  
W_q <- L_q / lambda

W <- W_q + 1 / mu

```


$W_q$ = `r W_q`

$W$ = `r W`

$L_q$ = `r L_q`

$L$ = `r L`

$\rho$ = `r rho`

Service time increased at the same times as the servers (3X), behaves almost the same as 1.


***



###2.6.12

```{r 2.6.12, echo=TRUE}

# Sign In 
m_int <- 6
m_serv <- 3
c <- 2
lambda <- 1/m_int
mu <- 1 / m_serv  

rho <- lambda / (c * mu) 

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho))) 

L_q <- round(((p0 * (lambda / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))),2)
L <- round((L_q + lambda / mu), 2) 
W_q <- round((L_q / lambda) ,2)
W <- round((W_q + 1 / mu), 2)

si <- c("Sign In", W_q, W, L_q, L, rho)

# Registration
m_serv <- 5
c <- 1
lambda_r <- 0.9 * lambda
mu <- 1 / m_serv  

rho <- lambda_r / (c * mu)

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho))) 
          
L_q <- round(((p0 * (lambda_r / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))),2)
L <- round((L_q + lambda_r / mu), 2) 
W_q <- round((L_q / lambda_r) ,2)
W <- round((W_q + 1 / mu), 2)

r <- c("Registration", W_q, W, L_q, L, rho)
  
# Trauma Rooms
m_serv <- 90
c <- 2
lambda_t <- 0.1 * lambda
mu <- 1 / m_serv  

rho <- lambda_t / (c * mu) 

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho))) 

L_q <- round(((p0 * (lambda_t / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))),2)
L <- round((L_q + lambda_t / mu), 2) 
W_q <- round((L_q / lambda_t) ,2)
W <- round((W_q + 1 / mu), 2)

tr <- c("Trauma Rooms", W_q, W, L_q, L, rho)

# Exam Rooms
m_serv <- 16
c <- 3
lambda_e <- 0.9 * lambda
mu <- 1 / m_serv  

rho <- lambda_e / (c * mu) 

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho))) 

L_q <- round(((p0 * (lambda_e / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))),2)
L <- round((L_q + lambda_e / mu), 2) 
W_q <- round((L_q / lambda_e) ,2)
W <- round((W_q + 1 / mu), 2)

er <- c("Exam Rooms", W_q, W, L_q, L, rho)

# Treatment Rooms
m_serv <- 15
c <- 2
lambda_t2 <- 0.9 * 0.6 * lambda + 0.1 * lambda
mu <- 1 / m_serv  

rho <- lambda_t2 / (c * mu) 

m <- c(0:(c-1))

p0 <- 1 / (sum((c * rho)^m / factorial(m)) + (c * rho)^c / (factorial(c) * (1- rho))) 

L_q <- round(((p0 * (lambda_t2 / mu)^c * rho) / ((factorial(c) * (1 - rho)^2))),2)
L <- round((L_q + lambda_t2 / mu), 2) 
W_q <- round((L_q / lambda_t2) ,2)
W <- round((W_q + 1 / mu), 2)

tr2 <- c("Treatment Rooms", W_q, W, L_q, L, rho)

hd <- c("Queue", "W-q", "W", "L_q", "L", "rho")

df = data.frame(hd, si, r, tr, er, tr2)
names(df) <- c(" ", "  ", "   ", "  ", "  ", "  ")
knitr::kable(df)
```


System is stable( $\rho$ < 1), Treatment room may use another server.